---
title: "MiBand Data Analysis"
author: "Borg"
date: "2016年3月16日"
output: html_document
---
# Overview
This report conducts data analysis of Xiaomi Mi Band's data from android sqlite database. 

# Data
For personalized purpose, the data can be obtained from android phone's path "/data/data/com.xiaomi.hm.health". Use RootExplorer to copy the whole directory named "databases" to your computer. Data used for this report can be accessed from [github](https://github.com/BigBorg/MiBand_Products)

# Load Libraries  
```{r message=FALSE}
if(!"MiBand" %in% installed.packages()){
  devtools::install_github('MiBand_R_Package','BigBorg')
}
library(MiBand)
library(ggplot2)
library(plotly)
```

# Load data
I've already packaged my code for data reading and cleaning inside MiBand package. If you are interested in the code inside the package, you can access them from my github repository.

```{r}
MiData <- loadMiData("./data/databases","963276123")
```

  
# Exploratory Data Analysis

```{r}
str(MiData)
head(MiData$data_clean)
```  
MiData is a list. Its element "data_clean" contains missing data, "data_week" is data with missing value substituted with mean value of the same day in week(mean value groupped by weekday).

```{r}
datedur<-range(MiData$data_week$date)
nrow<-nrow(MiData$data_week)
summary(MiData$data_week$sleep.light)
summary(MiData$data_week$sleep.deep)
summary(MiData$data_week$step)
```  
The date frame records data frome `r datedur[1]` to `r datedur[2]` and has `r nrow` rows. Sleep duration is recorded as count of minutes.  
  
Plotting:  
Histogram
```{r,message=FALSE,warning=FALSE}
ggplotly(miPlot(MiData,"hist","sleep"))
ggplotly(miPlot(MiData,"hist","step"))
```  


```{r message=FALSE,warning=FALSE}
ggplotly(miPlot(MiData,"box","sleep"))
ggplotly(miPlot(MiData,"ts","sleep"))
ggplotly(miPlot(MiData,"ts","step"))
```  

# Time Sequence Analysis  
Time sequence analysis on steps:
```{r}
weekly_ts_analysis <- function(data){
        tsobj <- ts(data,start=1,frequency=7)
        components <- decompose(tsobj)
        plot(components)
}
weekly_ts_analysis(MiData$data_week$step)
```  
  
Time sequence analysis on deep sleep:
```{r}
weekly_ts_analysis(MiData$data_week$sleep.deep)
```
  
Time sequence analysis on light sleep:
```{r}
weekly_ts_analysis(MiData$data_week$sleep.light)
```
  
# Which day of a week has the highest value?
```{r,warning=FALSE}
ggplotly(miPlot(MiData,"week","sleep"))
ggplotly(miPlot(MiData,"week","step"))
```  

# Is the step value of scool days different from that of vacation.
```{r}
MiData$data_week$month<-months(MiData$data_week$date)
vacation<-MiData$data_week[MiData$data_week$month %in% c("January","February","July","August"),]
schoolday<-MiData$data_week[!MiData$data_week$month %in% c("January","February","July","August"),]
boxplot(vacation$step,schoolday$step,names = c("vacation","school"))
title(main="Step")
```  
  
As shown in the boxplot, mean step of school day is higher than that of vacation.

```{r}
set.seed(0)
schoolresample<-matrix(sample(schoolday$step,1000,replace=T),nrow=100)
schoolmean<-apply(schoolresample,1,mean)
vacationresample<-matrix(sample(vacation$step,1000,replace = T),nrow = 100)
vacationmean<-apply(vacationresample,1,mean)
testresult<-t.test(schoolmean,vacationmean)
difference<-mean(schoolmean)-mean(vacationmean)
```
We are 1-`r testresult$p.value` confident to say step of school day is different from that of vacation. The mean difference is `r difference`(school Mean - Vacation Mean).

# Is there corelationship between sleep and step?

```{r}
MiData$data_week$efficiency<-with(MiData$data_week,sleep.deep/(sleep.deep+sleep.light))
cors<-with(MiData$data_week,c(
        cor(step,sleep.light),
        cor(step,sleep.deep),
        cor(step,sleep.light+sleep.deep),
        cor(step,efficiency)
        )
)
names(cors)<-c("step-sleep.light","step-sleep.deep","step-total sleep","step-efficiency")
cors
```  
Corelationship indicates that the longer you sleep, the fewer you are likely to walk. But such corelationship is quit weak. Note that within one row, which means on the same day, step of that day is recorded after sleep.

# At which point of total sleep you get highest sleep efficiency?
```{r message=FALSE}
# Use manipulate if you are copy-pasting code to R studio environment
# manipulate({
#        Y<-predict(loess(effciiency~I(sleep.light+sleep.deep),data=MiData$data_week),M)
#        ggplot(data=MiData$data_week,aes(sleep.light+sleep.deep,efficiency))+
#                geom_point()+geom_smooth(method="auto")+
#                geom_vline(x=M)+labs(x="Total sleep")+labs(title=paste("Efficiency: 
# ",Y,sep=""))
#    },
#    M=slider(
#            min(MiData$data_week$sleep.light+MiData$data_week$sleep.deep),
#            max(MiData$data_week$sleep.light+MiData$data_week$sleep.deep),
#            initial = min(MiData$data_week$sleep.light+MiData$data_week$sleep.deep)
#            )
#)
        ggplotly(ggplot(data=MiData$data_week,aes(sleep.light+sleep.deep,efficiency))+
                geom_point()+geom_smooth(method="auto")+
                labs(title="Efficiency"))
```  

The efficiency is extremely high when the total sleep is very small. That might be the body trying to compensate loss of total sleep time by increasing ratio of deep sleep. Though efficiency is high when you sleep for short time, deep sleep duration is not sufficent. As total sleep increase, we see a local high efficiency. Then efficiency goes down when you sleep for too long.

# Use total light sleep and deep sleep to predict step
```{r}
   coefs<-summary(lm(step~sleep.light+sleep.deep,data=MiData$data_week))$coefficient
```  
  
With deep sleep fixed, one minute increase of light sleep leads to `r coefs[2,1]` change of step. With light sleep fixed, one minute increase of deep sleep leads to `r coefs[3,1]` change of step.
  
# Conclusion
The subject sleep longer on Wednesday and walk more on Monday. Step of school day is different from that of vacation. There is a weak corelation between sleep and step. Around 7 hours' sleep has the relative high efficiency of sleep(deep sleep/total sleep). 