---
title: "MiBand Data Analysis"
author: "Borg"
date: "2016年3月16日"
output: html_document
---
# Overview
This report conducts data analysis of Xiaomi Mi Band's data from android sqlite database. 

# Data
For personalized purpose, the data can be obtained from android phone's path "/data/data/com.xiaomi.hm.health". Note that you need root access to copy files in the directory. Use RootExplorer to copy the whole directory named "databases" to R working directory. Data used for this report can be accessed from [github](https://github.com/BigBorg/MiBand)

# Load Libraries  
```{r message=FALSE}
library(DBI)
library(RSQLite)
library(jsonlite)
library(ggplot2)
library(gridExtra)
library(dplyr)
```

# Read Database File  
Using SQLite library to read in sqlite database file. Replace userid if you are running code to analyze your own data. You can find userid in Xiaomi Mi Band app.
```{r}
setwd("./databases")
userid<-"963276123"
filename<-paste("origin_db",userid,sep = "_")
sqlite    <- dbDriver("SQLite")
sourcedb <- dbConnect(sqlite,filename)
sourceData <- dbGetQuery(sourcedb,"select * from date_data")
setwd("..")
```
  
# Data Cleaning
Data is written in json form, thus using fromJSON and "[[" function to extract data from json. To make things clearer, bind useful date(date, light sleep, deep sleep and steps) together as a data frame called "cleanData."
```{r}
sourceData$DATE<-as.Date(sourceData$DATE,"%Y-%m-%d")
jsondata<-sapply(sourceData$SUMMARY,fromJSON)
slp<-jsondata[1,]
stp<-jsondata[4,]
lightsleep<-unname( unlist( lapply(slp,"[[","lt") )  )   ##extract elements from list of list
deepsleep<-unname( unlist( lapply(slp,"[[","dp") )  )
totalstep<-unname( unlist( lapply(stp,"[[","ttl") ) )
cleanData <- data.frame(date=sourceData$DATE,sleep.light=lightsleep,sleep.deep=deepsleep,step=totalstep)
cleanData<-dplyr::arrange(cleanData,date)
```  

## Missing Data Replacement  
Filling in missing value with mean value for the same day of a week.
```{r}
cleanData$weekday<-weekdays(cleanData$date)
groupeddata<-group_by(cleanData,weekday)
avgs<-summarize(groupeddata,
                sleep.light=mean(sleep.light[sleep.light!=0]),
                sleep.deep=mean(sleep.deep[sleep.deep!=0]),
                step=mean(step[step!=0]))
replace <- function(rowunit,avgs){
        for( col in 1:ncol(rowunit) ){
                if(rowunit[,col]==0){
                        rowunit[,col]<-avgs[avgs[,1]==rowunit$weekday,col]
                }
        }
        rowunit
}
completedata<-by(cleanData,1:nrow(cleanData),function(rowunit) replace(rowunit,avgs))
completedata<-as.data.frame(do.call("rbind",completedata))
```
  
# Exploratory Data Analysis

```{r}
head(completedata)
```  
The data frame has four columns called date, sleep.light, sleep.deep and step. The unit of time duration is minute.

```{r}
datedur<-range(completedata$date)
nrow<-nrow(completedata)
summary(completedata$sleep.light)
summary(completedata$sleep.deep)
summary(completedata$step)
```  
The date frame records data frome `r datedur[1]` to `r datedur[2]` and has `r nrow` rows.  
  
Plotting:
```{r message=FALSE}
boxplot(completedata$sleep.light/60,completedata$sleep.deep/60,names = c("light sleep","deep sleep"))
title("Sleep Box Plot")
ggplot(data=completedata,aes(x=date,y=value,col=type))+
        geom_point(aes(y=sleep.light/60,col="light sleep"))+
        geom_point(aes(y=sleep.deep/60,col="deep sleep"))+
        #geom_smooth(aes(y=sleep.light/60,col="light sleep"))+
        #geom_smooth(aes(y=sleep.deep/60,col="deep sleep"))+
        labs(y="sleep(hour)")
        
ggplot(data=completedata,aes(x=date,y=step))+
        geom_point()+
        #geom_smooth(method="auto")+
        labs(title="Steps")
```  

# Time Sequence Analysis  
Time sequence analysis on steps:
```{r}
weekly_ts_analysis <- function(data){
        tsobj <- ts(data,start=1,frequency=7)
        components <- decompose(tsobj)
        plot(components)
}
weekly_ts_analysis(completedata$step)
```  
  
Time sequence analysis on deep sleep:
```{r}
weekly_ts_analysis(completedata$sleep.deep)
```
  
Time sequence analysis on light sleep:
```{r}
weekly_ts_analysis(completedata$sleep.light)
```
  
# Which day of a week has the highest value?
```{r}
groupeddata<-group_by(completedata,weekday)
#avgs<-summarize(groupeddata,
#                sleep.light=mean(sleep.light[sleep.light!=0]),
#                sleep.deep=mean(sleep.deep[sleep.deep!=0]),
#                step=mean(step[step!=0]))
avgs<-avgs[c(4,2,6,7,5,1,3),] #sort by weekday

p1<-ggplot(data=avgs,aes(x=weekday,y=sleep/60,col="type"))+
        geom_line(aes(y=sleep.light/60,group=1,col="light sleep"))+
        geom_line(aes(y=sleep.deep/60,group=1,col="deep sleep"))+
        labs(title="Sleep Averages")+
        labs(y="average sleep(hour)")+
        theme(axis.text.x = element_text(face="bold", color="#993333",angle=45))
p2<-ggplot(data=avgs,aes(x=weekday,y=step))+
        geom_line(aes(group=1))+
        labs(title="Step Averages")+
        theme(axis.text.x = element_text(face="bold", color="#993333",angle=45))
grid.arrange(p1,p2,ncol=2)
maxindex<-data.frame()
for(col in 2:ncol(avgs)){
        if(col=="sleep.light") 
                maxindex<-t(avgs[,col]==max(avgs[,col]))
        maxindex<-rbind(maxindex,t(avgs[,col]==max(avgs[,col])))
}
max<-c()
for(row in 1:nrow(maxindex)){
        max<-c(max,avgs[as.numeric(maxindex[row,])==1,1])
}
names(max)<-rownames(maxindex)
```  
In the order of `r rownames(maxindex)`, `r max` have the highest value.

# Is the step value of scool days different from that of vacation.
```{r}
completedata$month<-months(completedata$date)
vacation<-completedata[completedata$month %in% c("January","February","July","August"),]
schoolday<-completedata[!completedata$month %in% c("January","February","July","August"),]
boxplot(vacation$step,schoolday$step,names = c("vacation","school"))
title(main="Step")
```  
  
As shown in the boxplot, mean step of school day is higher than that of vacation.

```{r}
set.seed(0)
schoolresample<-matrix(sample(schoolday$step,1000,replace=T),nrow=100)
schoolmean<-apply(schoolresample,1,mean)
vacationresample<-matrix(sample(vacation$step,1000,replace = T),nrow = 100)
vacationmean<-apply(vacationresample,1,mean)
testresult<-t.test(schoolmean,vacationmean)
difference<-mean(schoolmean)-mean(vacationmean)
```
We are 1-`r testresult$p.value` confident to say step of school day is different from that of vacation. The mean difference is `r difference`(school Mean - Vacation Mean).

# Is there corelationship between sleep and step?

```{r}
completedata$efficiency<-with(completedata,sleep.deep/(sleep.deep+sleep.light))
cors<-with(completedata,c(
        cor(step,sleep.light),
        cor(step,sleep.deep),
        cor(step,sleep.light+sleep.deep),
        cor(step,efficiency)
        )
)
names(cors)<-c("step-sleep.light","step-sleep.deep","step-total sleep","step-efficiency")
cors
```  
Corelationship indicates that the longer you sleep, the fewer you are likely to walk. But such corelationship is quit weak. Note that within one row, which means on the same day, step of that day happens after sleep.

# The longer you sleep, the higher efficiency of sleep you have?
```{r}
with(completedata,cor(sleep.deep+sleep.light,efficiency))
```
The longer you sleep, the lower efficiency of sleep you have.

# At which point of total sleep you get highest sleep efficiency?
```{r message=FALSE}
# Use manipulate if you are copy-pasting code to R studio environment
# manipulate({
#        Y<-predict(loess(effciiency~I(sleep.light+sleep.deep),data=completedata),M)
#        ggplot(data=completedata,aes(sleep.light+sleep.deep,efficiency))+
#                geom_point()+geom_smooth(method="auto")+
#                geom_vline(x=M)+labs(x="Total sleep")+labs(title=paste("Efficiency: 
# ",Y,sep=""))
#    },
#    M=slider(
#            min(completedata$sleep.light+completedata$sleep.deep),
#            max(completedata$sleep.light+completedata$sleep.deep),
#            initial = min(completedata$sleep.light+completedata$sleep.deep)
#            )
#)
Y<-predict(loess(efficiency~I(sleep.light+sleep.deep),data=completedata),415)
        ggplot(data=completedata,aes(sleep.light+sleep.deep,efficiency))+
                geom_point()+geom_smooth(method="auto")+
                geom_vline(xintercept=415)+labs(x="Total sleep")+
                labs(title=paste("Efficiency: ",Y,sep=""))
```  

The efficiency is extremely high when the total sleep is less than 300. That might be the body trying to compensate loss of total sleep time by increasing deep sleep percentage. Though efficiency is high when you sleep for less than 300 minutes, deep sleep duration is not sufficent for the total sleep duration is short. At about 415 minutes, we see a slightly high efficiency relatively. Then efficiency goes downward again when you sleep for longer than that.

# Relationship between total sleep duration and deep sleep duration
```{r}
ggplot(data=completedata,aes(x=sleep.light+sleep.deep,y=sleep.deep))+
                geom_point()+
                geom_smooth(method="auto")+
                labs(title="Deep Sleep/Total Sleep",x="Total Sleep(minute)",y="Deep Sleep(minute)")
```  
Again, using deep sleep duration to evaluate sleep quality, the turning point is still at about 415 minutes.

# Use total light sleep and deep sleep to predict step
```{r}
   coefs<-summary(lm(step~sleep.light+sleep.deep,data=completedata))$coefficient
```  
  
One minute increase of light sleep leads to `r coefs[2,1]` change of step. One minute increase of deep sleep leads to `r coefs[3,1]` change of step.
  
# Conclusion
The subject sleep longer on Sunday and walk more on Monday. Step of school day is different from that of vacation. There is a weak corelation between sleep and step. Around 7 hours' sleep has the relative high efficiency of sleep(deep sleep/total sleep). 